DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-20 paper/intro.tex
intro revision

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability with
no human intervention. Deep neural networks (DNNs) have been
successfully applied in various perception and control tasks in
recent years.  They are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which used a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNN based Artificial Intelligence workloads may be used in future
autonomous vehicles.

@@

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely deep learning based end-to-end real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN, which runs on a significantly more powerful computing
platform (Drive PX~\cite{drivepx}).

@@

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of deep-learning based real-time
control, especially on real-time deep neural network \emph{inferencing}.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.
From the systematic study, we want to answer the following questions:
(1) how well does the DNN based real-time inferencing task perform on
contemporary embedded multicore computing platforms? (2) how
susceptible is the DNN inferencing task to contention in shared
hardware resources (e.g., cache and DRAM) when multiple tasks/models
are consolidated? (3) are the existing state-of-the-art shared resource
isolation techniques effective in protecting real-time performance of
the DNN inferencing task?

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
Our observations are as follows.
First, we find that real-time processing of the DNN inferencing is
feasible on contemporary embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. Second, while consolidating
multiple DNN models and task are feasible but the impact of shared
resource contention can be catastrophic in the worst-case as we
observe up to 11.6X slowdown even when the DNN model was given a
dedicated core. Third, using the state-of-the-art OS-level cache
partitioning and memory bandwidth control
mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
minimal performance impact to the DNN workload while memory
performance has a significant impact.

@@

% contributions
This paper makes the following {\bf contributions}:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    employs a state-of-the-art CNN based end-to-end
    real-time control.~\footnote{The source code, datasets, and build
      instructions of DeepPicar can be found at:
      \url{https://github.com/mbechtel2/DeepPicar-v2}}.
  \item We provide extensive empirical performance evaluation results
    of the CNN inferencing workload on a number of contemporary
    embedded computing platforms.
  \item We apply the state-of-the-art shared resource isolation
    techniques and evaluate their effectiveness in protecting the
    inferencing workload in consolidated settings.
\end{itemize}

The remaining sections of the paper are as follows.
Section~\ref{sec:background}
provides a background of applications in autonomous driving and related works.
Section~\ref{sec:overview} gives an overview of the DeepPicar
platform, including the high-level system and the methods used for
training and inference. Section~\ref{sec:evaluation} presents our
evaluation of the platform and how different factors can affect
performance. Section~\ref{sec:comparison} offers a comparison between  
the Raspberry Pi 3 and other embedded computing platforms to 
determine their suitably for autonomous driving research.
%% We discuss more detailed training methods and system-level issues
%% in~\ref{sec:discussion}. 
We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.

>>>>

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability with
no human intervention. Deep neural networks (DNNs) have been
successfully applied in various perception and control tasks in
recent years.  They are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which used a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNN based artificial intelligence (AI) workloads may be used in future
autonomous vehicles.

@@

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely end-to-end deep learning based real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN.
%% Note that, DAVE-2 run on a significantly more powerful computing
%% platform (Drive PX~\cite{drivepx}).

@@

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of end-to-end deep-learning based real-time
control, especially on real-time \emph{inferencing} of the CNN.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.
From the systematic study, we want to answer the following questions:
(1) How well does the CNN based real-time inferencing task perform on
contemporary embedded multicore computing platforms? (2) How
susceptible is the CNN inferencing task to contention in shared
hardware resources (e.g., cache and DRAM) when multiple tasks/models
are consolidated? (3) Are the existing state-of-the-art shared resource
isolation techniques effective in protecting real-time performance of
the CNN inferencing task?

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
Our observations are as follows.
First, we find that real-time processing of the CNN inferencing is
feasible on contemporary embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. Second, while consolidating
multiple CNN models and task are feasible but the impact of shared
resource contention can be catastrophic in the worst-case as we
observe up to 11.6X slowdown even when the CNN model was given a
dedicated core. Third, using the state-of-the-art OS-level cache
partitioning and memory bandwidth control
mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
minimal performance impact to the CNN workload while memory
performance has a significant impact.

@@

% contributions
This paper makes the following {\bf contributions}:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    employs a state-of-the-art CNN based end-to-end
    real-time control~\footnote{The source code, datasets, and build
      instructions of DeepPicar can be found at:
      \url{https://github.com/mbechtel2/DeepPicar-v2}}.
  \item We provide extensive empirical performance evaluation results
    of the CNN inferencing workload on a number of contemporary
    embedded computing platforms.
  \item We apply the state-of-the-art shared resource isolation
    techniques and evaluate their effectiveness in protecting the
    inferencing workload in consolidated settings.
\end{itemize}

The remainder of the paper is organized as follows.
Section~\ref{sec:background} provides a background on the application
of nueral networks in autonomous driving.
Section~\ref{sec:overview} gives an overview of the DeepPicar
testbed. Section~\ref{sec:evaluation} presents extensive real-time
performance evaluation results we collected on the testbed.
Section~\ref{sec:comparison} offers a comparison between  
the Raspberry Pi 3 and other embedded computing platforms.
%% We discuss more detailed training methods and system-level issues
%% in~\ref{sec:discussion}. 
We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.
